---
title: "Data Manipulation for NHIRD"
author: "Nick Lin"
date: "6/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Fundamental

## Version and setting

First for all
```{r}
sessionInfo()
```

設置成 UTF-8
- Tools > Global Options > Code > Saving > Default text encoding > UTF-8


## Rstudio environment

- Four panels
  + Script
  + Console
  + Environment, History, Connections, Tuttorial
  + Files, Plots, Packages, Help, Viewer
- R project
- Rmd
  + Reference
  + Knit

## Data structure

Vector
- Atomic vector
  + Logical
  + Integer (factor, date) \in numeric(number)
  + Double (time) \in numeric(number)
  + Character
  + (vector, matrix, array)
- List (data.frame, tibble)

command+option+i / cntl+alt+i
```{r}
1+1
a = c(1L, 2L, 3L) #integer
a
typeof(a)
mode(a)

d = c(3.3, 3.14)
d
typeof(d) # vvvv
mode(d) # xxxx

b = c("ha", 3, "hello") #charcter
b
typeof(b)

l = list("a", 3L, 3.24)
l
```

## How to use help

```{r}
help(xxxx)
?xxxx
??xxxx
```


# Load and write data

## Install related packages

Please update to the newest version.
```{r}
# needPackages <- c("tidyverse")
# install.packages(needPackages)

library(tidyverse) # for data manipulation (https://www.tidyverse.org/)
library(haven) # for read sas, spss file (https://github.com/tidyverse/haven)
```

## Get directions and files

### read SAS code

```{r}
dir("rawdata/sas/")
data_dirs <- list.dirs(path = "rawdata/sas", recursive = FALSE)
data_dirs
data_dirs[1]
health01_files <- list.files(path = data_dirs[1], full.names = TRUE)
health01_files
health01_files[2]
```


haven::read_sas()
```{r}
d1 <- read_sas(health01_files[1], n_max = 5)
names(d1)
dim(d1)
d1

d2 <- read_sas(health01_files[1], skip = 5, n_max = 5)
d2

d3 <- read_sas(health01_files[1]) # n_max = inf
d3
```


```{r}
d1 <- read_sas(health01_files[1], 
               col_select = HOSP_ID:SEQ_NO)
d1

d2 <- read_sas(health01_files[1], 
               col_select = c(HOSP_ID:APPL_DATE, DRUG_DOT, CHR_DAYS),
               n_max = 10)
d2


a = c(1, 4, 9)
print(sum(sqrt(sqrt(a))))
a %>% sqrt() %>% sqrt() %>% sum() %>% print()

d2_changeType <- d2 %>% 
  mutate(FEE_YM = zoo::as.yearmon(FEE_YM, format = "%Y%m", tryformat), 
         APPL_TYPE = as.integer(APPL_TYPE),
         DRUG_DOT = as.character(DRUG_DOT))
d2_changeType

# d2_changeType <- mutate(d2, ......)

d2
```
POXIClt / POXICct

Date 2020/02/03
Time 12:15:30
DateTime 2020/02/03 12:15:30

```{r}
d3

write_csv(d3, "data/test1.csv")

d33 <- read_csv("data/test1.csv")
d33
```





### read CSV data

```{r}
dir("rawdata/csv/")
data_dirs <- list.dirs(path = "rawdata/csv/", recursive = FALSE)
data_dirs
health01_files <- list.files(path = data_dirs[1], full.names = TRUE)
health01_files
health01_files[1]
```

```{r}

df1 <- read_csv(health01_files[1])
df1

head(df1, 1) # head(df1, 10)
tail(df1)
View(df1)

df2 <- read_csv(health01_files[2])
df2

df3 <- read_csv(health01_files[3])
```


## Read subset of data
```{r}
.col_type <- cols_only(HOSP_ID = col_character(),
                       ICD9CM_2 = col_character(), 
                       FEE_YM = col_character())

df1 <- read_csv(health01_files[5], col_types = .col_type,
                n_max = 10)
df1
```
```{r}
write_csv(df1, "data/health01_new.csv")
df11 <- read_csv("data/health01_new.csv")
df11
```



## 0627_Exercise

0. sessionInfo()
1. 更新所需要的軟體與套件，確認版本是最新的
2. 整理好你的工作資料夾，包含
  - rawdata
    - csv
      - Health-01
      - Health-02
      - ...
    - sas
      - Health-01
      - Health-02
      - ...
  - data 
3. 嘗試讀清楚上方所用到的程式碼，試著改改看任一個參數會有什麼變化
  - 例如`[i]`裡頭的`i`改成不同的數值
  - 
4. 選出你每筆 data 所需要的 variables (cols) 以及對照資料庫使用手冊所規定的變數型態
  - character, integer, double, logical, factor, date, ...

例如：

health01
- HOSP_ID = col_character()
- ICD9CM_2 = col_character()
...

health02


health03

5. 有空的話看一下 Rmarkdown 的介紹：https://rmarkdown.rstudio.com/lesson-1.html



## Combind data in the same folder


for loop

```{r}
for (i in 1:10) {
  print(i)
}
```


```{r}
n_files <- 2
# n_files <- length(health01_list)
health01_list <- vector(mode = "list", length = n_files)

for (i in 1:n_files) {
  health01_list[[i]] <- read_csv(health01_files[i], col_types = .col_type)  
}

health01_df <- bind_rows(health01_list)
health01_df
```

### Write csv

```{r}
write_csv(health01_df, file = "data/health01.csv")
```



# Data manipulation

Reference: []()

## data.frame and tibble

## create, select, subset, remove ...

## dplyr



